---
title: "Dashboard Demo"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
---

```{r global, include=FALSE}
options(tigris_use_cache = TRUE)
library(flexdashboard)
library(tidyverse)
library(leaflet)
library(sf)
library(plotly)
library(tigris)

sensor_data_all_time <-
readRDS("sensor_data_all_time.rds")

sensor_data_ID_map_AQI_transformed <-
readRDS("sensor_data_ID_map_AQI_transformed.rds")

smc_pm25 <- readRDS("smc_pm25.rds")

smc_race_income_leftlet <-
readRDS("smc_race_income_leftlet.rds")

smc_race_income_leftlet_Top <-
readRDS("smc_race_income_leftlet_Top.rds")

```

Page 1
================================

Inputs {.sidebar}
-------------------------------------

```{r}
checkboxGroupInput(
  inputId = "CITY",
  label = "Jurisdiction: ",
  choices = c("Redwood City","Palo Alto"),
  selected = "Redwood City",
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)

selectInput(
  inputId = "LOCATION",
  label = "LOCATION: ",
  choices = c("inside","outside"),
  selected = "outside"
)

```

Column
-------------------------------------
### PART I. PM25 Exposure in Februnary, 2022
```{r}
plotlyOutput("plot")
```

```{r, context = "server"}
observeEvent({
  input$LOCATION
  input$CITY
  
}, {
  
    chart <- sensor_data_all_time %>% 
    filter(
      Location %in% c(input$LOCATION),
      city %in% c(input$CITY)
    ) %>% 
  ggplot() +
  geom_line(
    aes(
      x = date,
      y = PM25,
      color=city
     )
  ) +
  labs(
      x = "Date",
      y = "PM2.5",
      title = "PM2.5 exposure"
    ) 
  
   output$plot <- renderPlotly({
    chart %>% 
      ggplotly() %>% 
      config(displayModeBar = F)
  })
  
})
```

Column
-------------------------------------

### PART I. AQI in Februnary, 2022

```{r}
leafletOutput("map1")
```

```{r}
observeEvent({
  input$LOCATION
  input$CITY
}, {

   AQI_sensor <-
    sensor_data_ID_map_AQI_transformed %>% 
    filter(
      Location == input$LOCATION,
      city == input$CITY
    ) 
  
  aqi_pal1 <- colorQuantile(
  palette = "RdYlGn",
  reverse = T,
  domain =  AQI_sensor$AQI, n = 5
    )

 output$map1 <- renderLeaflet({
    AQI_sensor %>% 
    leaflet() %>% 
  addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addCircleMarkers(
    color = ~aqi_pal1(AQI),
    label = ~paste0(AQI,", ",AQI_Cat),
    radius = 5,
    opacity = 0.75
  ) %>% 
  addLegend(
    pal = aqi_pal1,
    values = ~AQI
  )

  })
  
})
```

Page 2
================================

Inputs {.sidebar}
-------------------------------------

```{r}
checkboxGroupInput(
  inputId = "CITY2",
  label = "Jurisdiction: ",
  choices = c("Redwood City","Palo Alto"),
  selected = "Redwood City",
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)

selectInput(
  inputId = "LOCATION2",
  label = "LOCATION: ",
  choices = c("inside","outside"),
  selected = "outside"
)

```

Column
-------------------------------------

###  PART I. AQI in Februnary, 2022

```{r}
leafletOutput("map2")
```

```{r}
observeEvent({
  input$LOCATION2
  input$CITY2
}, {
   
    AQI_sensor <-
    sensor_data_ID_map_AQI_transformed %>% 
    filter(
      Location == input$LOCATION2,
      city == input$CITY2
    ) 
  
  aqi_pal2 <- colorFactor(
  palette = "RdYlGn",
  reverse = T,
  domain = sensor_data_ID_map_AQI_transformed$AQI_Cat
)

 output$map2 <- renderLeaflet({
    AQI_sensor %>% 
    leaflet() %>%  
   addProviderTiles(provider = providers$CartoDB.Positron) %>% 
  addCircleMarkers(
    color = ~aqi_pal2(AQI_Cat),
    label = ~AQI_Cat,
    radius = 5,
    opacity = 0.75
  ) %>% 
  addLegend(
    pal = aqi_pal2,
    values = ~AQI_Cat
  )

  })
  
})   
   
```
   
Page 3
================================

Inputs {.sidebar}
-------------------------------------

```{r}
checkboxGroupInput(
  inputId = "TYPE",
  label = "TYPE: ",
  choices = c("Income","Race"),
  selected = "Income",
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)

selectInput(
  inputId = "LOCATION3",
  label = "LOCATION: ",
  choices = c("inside","outside"),
  selected = "outside"
)

```


Column 
---------------------------------
### PART II. PM25 Exposure by Race/Income
```{r}
plotlyOutput("plot3A")
```

```{r, context = "server"}
observeEvent({
  input$LOCATION3
  input$TYPE 
}, {
  
    data1 <- smc_pm25 %>% 
    filter(
       Location %in% c(input$LOCATION3),
       type %in% c(input$TYPE)
    )  
      
    chart1 <- data1 %>% 
    ggplot() +
    geom_bar(
    aes(x = PM25_tier %>% factor
                   (levels = c("0-2", "2-4",
                              "4-6", "6-8",
                              "8-10", "10-12",
                              "12-14", "14+" )),
        y =  estimate,
        fill = group 
       ),
        stat = "identity",
        position = "stack"
  ) + 
labs(
    x = "PM2.5",
    y = "Number of households",
    fill="Group",
    title = "PM2.5 exposure in San Mateo County"
  ) + 
  coord_flip()
    theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  ) 
        
  
   output$plot3A <- renderPlotly({
    chart1 %>% 
      ggplotly() %>% 
      config(displayModeBar = F)
  })
  
})
```

Column
-------------------------------------
### PART II. PM25 Exposure - Equity Analysis
```{r}
plotlyOutput("plot3B")
```

```{r, context = "server"}
observeEvent({
  input$LOCATION3
  input$TYPE
  
}, {
  
    data2 <- smc_pm25 %>% 
      filter(
      Location %in% c(input$LOCATION3),
      type %in% c(input$TYPE)
    )  
      
    chart2 <- data2 %>% 
    ggplot() +  
    geom_bar(
    aes(x = PM25_tier %>% factor
                   (levels = c("0-2", "2-4",
                              "4-6", "6-8",
                              "8-10", "10-12",
                              "12-14", "14+" )),
        y =  estimate,
        fill = group 
       ),
        stat = "identity",
        position = "fill"
  ) + 
labs(
    x = "PM2.5",
    y = "Proportion of households",
    fill="Group",
    title = "PM2.5 exposure in San Mateo County"
  ) + 
  coord_flip()
    theme(
    legend.position = "bottom",
    legend.direction = "vertical"
  )  +
  guides(
    fill = guide_legend(
      reverse = T
    )
  ) 
        
  
   output$plot3B <- renderPlotly({
    chart2 %>% 
      ggplotly() %>% 
      config(displayModeBar = F)
  })
  
})
```

Page 4
================================

Inputs {.sidebar}
-------------------------------------

```{r}
checkboxGroupInput(
  inputId = "LOCATION4",
  label = "LOCATION: ",
  choices = c("inside","outside"),
  selected = "outside", 
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)

```


Column 
---------------------------------
### PART III. Suggested equity scores based on sensor count, white population, $75K-$100K population in San Mateo County

```{r}
leafletOutput("map4")
```

```{r, context = "server"}
observeEvent({
  input$LOCATION4
}, {

   sensor_placing <- 
    smc_race_income_leftlet %>% 
    filter(
      Location == input$LOCATION4
    ) 
  
   res_pal <- colorNumeric(
   palette = "RdYlGn",
   domain = sensor_placing$equity_score
   )

   output$map4 <- renderLeaflet({
   leaflet() %>%
   addTiles() %>%
   addPolygons(data = sensor_placing,
   fillColor = ~res_pal(equity_score),
   color = "white",
   opacity = 0.5,
   fillOpacity = 0.5,
   weight = 1,
   label = "Sensor placing priority",
   highlightOptions = highlightOptions(
   weight = 2,
   opacity = 1
   ) ) %>%
   addLegend(data = sensor_placing,
   pal = res_pal,
   values = ~equity_score,
   title = "Suggested equity scores")

     })
  
})

```
   
   
Page 5
================================

Inputs {.sidebar}
-------------------------------------

```{r}
checkboxGroupInput(
  inputId = "LOCATION5",
  label = "LOCATION: ",
  choices = c("inside","outside"),
  selected = "outside", 
  inline = FALSE,
  width = NULL,
  choiceNames = NULL,
  choiceValues = NULL
)

selectInput(
  inputId = "RANK",
  label = "TOP PRIORITY: ",
  choices = c("1","5","10","20","30","40","50"),
  selected = "1"
)

```


Column 
---------------------------------
### PART III. Where to place the sensor according to the priority from the recommended equity scores in San Matero County

```{r}
leafletOutput("map5")
```

```{r, context = "server"}
observeEvent({
  input$RANK
  input$LOCATION5
}, {

   data_placing <- 
    smc_race_income_leftlet_Top %>% 
    filter(
      Location == input$LOCATION5,
      Rank == input$RANK
    ) 
  
   res_pal <- colorNumeric(
   palette = "RdYlGn",
   domain = data_placing$Top
   )

   output$map5 <- renderLeaflet({
   leaflet() %>%
   addTiles() %>%
   addPolygons(data = data_placing,
   fillColor = ~res_pal(Top),
   color = "white",
   opacity = 0.5,
   fillOpacity = 0.5,
   weight = 1,
   label = "Recommended sensor locations",
   highlightOptions = highlightOptions(
   weight = 2,
   opacity = 1
   ) ) %>%
   addLegend(data = data_placing,
   pal = res_pal,
   values = ~Top,
   title = "Sensor placing locations<br>with top priorities")

     })
  
})

```
   



