---
title: "Assignment 3"
author: "Lillian (Wei) Hung"
date: "2021/10/10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```



#  Bay_counties

```{r}
library(devtools)
# install_github('walkerke/tigris', force = TRUE )

Sys.setenv(CENSUS_KEY="c8100bd1d2b42c5942616b53158bee261f69edbe")

```

```{r}
library(censusapi)
library(tidyverse)
library(tigris)
library(sf)
library(leaflet)
library(mapview)
```

```{r}

ca_counties <- counties("CA", cb = T, progress_bar = F)

bay_county_names <-
  c(
    "Alameda",
    "Contra Costa",
    "Marin",
    "Napa",
    "San Francisco",
    "San Mateo",
    "Santa Clara",
    "Solano",
    "Sonoma"
  )

bay_counties <-
  counties("CA", cb = T, progress_bar = F) %>%
  filter(NAME %in% bay_county_names)

```



```{r}
Sys.setenv(CENSUS_KEY="c8100bd1d2b42c5942616b53158bee261f69edbe")

pums_2019_1yr <- getCensus(
  name = "acs/acs1/pums",
  vintage = 2019,
  region = "public use microdata area:*", 
  regionin = "state:06",
  vars = c(
    "SERIALNO",
    "SPORDER",
    "PWGTP",
    "WGTP",
    "NP",
    "HHLANP",
    "HHL",
    "HINCP",
    "FPARC",
    "HUGCL",
    "AGEP",
    "SCHL",
    "TEN"
  )
)

```


# --------------------------------

```{r}

ca_pumas <-
  pumas("CA", cb = T, progress_bar = F)

bay_pumas <-
  ca_pumas %>% 
  st_centroid() %>% 
  .[bay_counties, ] %>% 
  st_drop_geometry() %>% 
  left_join(ca_pumas %>% select(GEOID10)) %>% 
  st_as_sf()

bay_pums <-
  pums_2019_1yr %>% 
  mutate(
    PUMA = str_pad(public_use_microdata_area,5,"left","0")
  ) %>% 
  filter(PUMA %in% bay_pumas$PUMACE10)

# -------------------------------

bay_pums_A3 <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(
    WGTP = as.numeric(WGTP),
    heat_impact_pop = ifelse(
        (HINCP < 22000) &
        (AGEP<6|AGEP>64) &
        (TEN=3) & (HHL != 1) & (SCHL <16),
    WGTP,
      0
    )
  ) %>% 
  group_by(PUMA) %>% 
  summarize(
    perc_heat_impact_pop =
      sum(heat_impact_pop, na.rm =T)/sum(WGTP, na.rm = T)*100
  ) %>% 
  left_join(
    bay_pumas %>% 
      select(PUMACE10),
    by = c("PUMA" = "PUMACE10")
  ) %>% 
  st_as_sf()


pums_pal <- colorNumeric(
  palette = "orange",
  domain = bay_pums_A3$perc_heat_impact_pop
)

leaflet() %>%
  addTiles() %>% 
  addPolygons(
    data = bay_pums_A3,
    fillColor = ~pums_pal(perc_heat_impact_pop),
    color = "green",
    opacity = 0.5,
    fillOpacity = 0.5,
    weight = 1,
    label = ~paste0(
      round(perc_heat_impact_pop), 
      "%"
    ),
      highlightOptions = highlightOptions(
      weight = 2,
      opacity = 1
    )
  ) %>% 
  addLegend(
    data = bay_pums_A3,
    bins=5,
    pal = pums_pal,
    values = ~perc_heat_impact_pop,
    title = "% Households with children Under 5 or elderly,</br> for poverty, living in rented house,</br> less educated, and non-English speaking groups"
  )

```

```{r}
heat_impact_total_pop <-
  bay_pums %>% 
  filter(!duplicated(SERIALNO)) %>% 
  mutate(WGTP = as.numeric(WGTP),
    WGTP = as.numeric(WGTP),
    heat_impact_total_pop = ifelse(
        (HINCP < 22000) &
        (AGEP<6|AGEP>64) &
        (TEN=3) & (HHL != 1) & (SCHL <16),
    WGTP,
      0    
    )
  )

heat_impact_total_pop <-
summarize(
    heat_impact_total_pop,Total_heat_impact_population=
      sum(heat_impact_total_pop, na.rm =T)
  )

heat_impact_total_pop 

```

From the map, we can collect information on people with vulnerability characteristics and thus might need safety resources from institutions. According to the data of California heat assessment website, I chose the following vulnerability characteristics: “Households with children Under 5 or elderly, for poverty, living in a rented house, less educated, and non-English speaking groups” to produce a map. Researchers in California have been analyzing the impact of heat waves in recent years. Due to the extreme temperature of heat waves, some people in California suffered heat-related illnesses. Therefore, producing maps of this issue can be useful for us to prepare for heat waves. 

This map includes useful information about people who are vulnerable to heat waves. From the analysis of this map, there are 4%-5% of people with these vulnerability characteristics in the area of Richmond, Oakland, and Martinez. Additionally, there are 7% of people with these vulnerability characteristics in the area near Little Portugal. When heat wave hits these areas, it might affect people’s health. They might get dehydrated and send to the hospital. Thus, the government, public institutions, local organizations should be aware of this environmental issue and take actions to protect people who didn’t have access to the safety resources. For instance, they can offer food and water to support them. In addition, they can have a partnership with local clinics and hospitals.










